### 2021-04-09 introduction.tex
another pass to intro

With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and the $\pi$-calculus with binary session types~\cite{cairespfenning10,wadler14}, which guarantees deadlock freedom by restricting the allowed processes to only trees. This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks. Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs, either by maintaining a strong correspondence to classical linear logic, as in Hypersequent CP~\cite{kokkemontesi19popl,kokkemontesi19tlla}, or by weakening the correspondence to classical linear logic in exchange for a more expressive language, as in Priority CP (PCP)~\cite{dardhagay18}. PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule. Differently from previous work, PCP allows cyclic structures and restores deadlock freedom by adding priorities to types following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}. Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleaving of channels. Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof) by leveraging priorities. Then, deadlock freedom follows as a result of the cycle-elimination proof.
%
With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom. Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork. The fork construct is the term representation of the cut construct in CP, which as mentioned above, evaluates to a channel restriction combined with a parallel composition. As with CP, this is restrictive because it limits process structures to only trees.

>>>>

With regards to $\pi$-calculus, a growing line of work on deadlock freedom leverages the Curry-Howard correspondences between intuitionistic or classical linear logic and the $\pi$-calculus with binary session types~\cite{cairespfenning10,wadler14}, which guarantees deadlock freedom by restricting the allowed processes to only trees. This is achieved by merging constructs for parallel composition and channel restriction under the logical cut rule, which forces processes to share \emph{only one} channel for communication, thus eliminating cyclic dependencies and deadlocks. Recent developments in the Classical Processes (CP)~\cite{wadler14} line of work have led to various approaches to decoupling these constructs. One way is by maintaining a strong correspondence to classical linear logic and allowing only tree-structured processes, such as in Hypersequent CP (HCP)~\cite{kokkemontesi19popl,kokkemontesi19tlla}, Linear Compositional Choreographies (LCC)~\cite{CarboneMS18} and Classical Transitions (CT)~\cite{MP18}. Another way is by weakening the correspondence to classical linear logic in exchange for a more expressive language where cyclic-structured processes are allowed, such as in Priority CP (PCP)~\cite{dardhagay18}. PCP decouples CP's cut rule into two separate constructs: a construct for parallel composition, by introducing a mix rule; and a construct for restriction, by introducing a cycle rule. This decoupling allows cyclic processes. In order to restore deadlock freedom, PCP uses priorities following Kobayashi~\cite{kobayashi06} and Padovani~\cite{padovani14}. Priorities establish an order of actions in a $\pi$-calculus process and are used to rule out bad interleavings of channels. Dardha and Gay prove cycle-elimination of PCP (following the standard cut-elimination proof) by leveraging priorities. Then, deadlock freedom follows as a result of the cycle-elimination proof.
%
With regards to GV, the original work on session types in GV~\cite{gayvasconcelos12} did not satisfy deadlock freedom. Later on, calculi in the GV family~\cite{wadler15,lindleymorris15} have achieved deadlock freedom via a syntactic restriction, \ie by combining channel creation and thread spawning into a single operation, called fork. The fork construct is the term representation of the cut construct in CP, which as mentioned above, evaluates to a channel restriction combined with a parallel composition. As with CP, this is restrictive because it limits process structures to only trees.
